---
title: "hw2"
author: "Mao Xin"
output: html_document
---

## Import libarary

```{r}
if (!require("gplots")) {
  install.packages("gplots", dependencies = TRUE)
  library(gplots)
} # for the use of heatmap.2

if (!require("RColorBrewer")) {
  install.packages("RColorBrewer", dependencies = TRUE)
  library(RColorBrewer)
} # provide pallet for heatmap.2

if (!require("FactoMineR")) {
  install.packages("FactoMineR", dependencies = TRUE)
  library(FactoMineR)
} # advanced PCA
```

## Clustering without PCA

```{r}
setwd('/Users/maoxin/Desktop/bio_HomeworkII')
df <- read.delim("GeneMatrix.txt")
df_1 <- as.data.frame(t(df))

distance <- dist(df_1, method='euclidean')
m_dist <- as.matrix(distance)
cl_func <- function(foo) {
  hclust(foo, method = "average")
}
heatmap.2(m_dist, trace='none', labRow=NA, labCol=NA, 
          dendrogram="col", col=brewer.pal(9, "YlOrRd"), hclustfun=cl_func)
# heatmap with hclust with the 'average' method
# two cluster group can be detected

cl <- hclust(distance, method='average')
cl_result <- cutree(cl, k=2)
plot(cl_result)
# group 1 contains 413 observations and group 2 contains 109 ones
```

## Clustering with PCA (Threshold set to be 0.8)

```{r}
pca_result <- PCA(df_1, scale.unit=T, ncp=ncol(df_1), graph=F)
df_2 <- pca_result$ind$coord[,1:tail(which(pca_result$eig$`cumulative percentage of variance` <=80), n=1)]
# set the pca threshold to be 0.8

distance_2 <- dist(df_2, method='euclidean')
m_dist_2 <- as.matrix(distance_2)
heatmap.2(m_dist_2, trace='none', labRow=NA, labCol=NA, 
          dendrogram="col", col=brewer.pal(9, "YlOrRd"), hclustfun=cl_func)
# two cluster group can be detected

cl_2 <- hclust(distance_2, method='average')
cl_result_2 <- cutree(cl_2, k=2)
plot(cl_result_2)
# group 1 contains 431 observations and group 2 contains 91 ones
```

## Evaluate the performance based on artificial classification
```{r}
test_df <- read.delim("clinical_data")
test_df$sampleID = gsub("-", ".", test_df$sampleID)
rownames(test_df) <- test_df$sampleID
test_df_1 <- test_df[rownames(df_2),]
test_df_1$ER_Status_nature2012 <- gsub("Positive", 1, test_df_1$ER_Status_nature2012)
test_df_1$ER_Status_nature2012 <- gsub("Negative", 2, test_df_1$ER_Status_nature2012)
test_df_1$ER_Status_nature2012 <- as.numeric(test_df_1$ER_Status_nature2012)
plot(test_df_1$ER_Status_nature2012)
# group 1 (ER_Status_nature2012=='Positive') contains 404 observations and group 2 (ER_Status_nature2012=='Negative') contains 118 ones

error_direct <- sum(abs(test_df_1$ER_Status_nature2012 - cl_result)) / nrow(df_1)
error_pca <- sum(abs(test_df_1$ER_Status_nature2012 - cl_result_2)) / nrow(df_1)
# calculate the total number of wrong clustering

print(error_direct)
print(error_pca)
```

## Discussion

The direct clustering, PCA enhanced clustering, and artificial classification show the same result of two groups with the first one contaions about 400 observations.

The error rate of the direct clustering is 0.06321839 while the rate of PCA enhanced clustering increases to 0.09770115. This is caused by the omitting of information with weak correlation. 

