---
title: "homework2_2012011467"
author: "Shaoming Song"
date: "2015年12月7日"
output: html_document
---
##Hierarchical Clustering Based on original data
##Special thanks to Rhodes Zhang for teaching me how to use R ^_^

```{r}

library(RColorBrewer)
library(gplots)

##load data
setwd("C:/study/大四上/生物信息学/homework2")
gene_expression <- read.table(file = "GeneMatrix.txt")
gene_expression <- t(gene_expression)
gene_expression <- as.data.frame(gene_expression)

##get length & distance to clust
length <- length(gene_expression[1,])
d <- dist(gene_expression)

##Hierarchical Clustering & get the results of classification
hc <- hclust(d , method = "average")
model <- hc
result=cutree(model,k=2)

##define a function which is used in drawing heatmap
myclust <- function(x){
  hclust(x , method = "average")
}
```
##Heatmap1
```{r echo=FALSE}
heatmap.2(as.matrix(d),
          trace='none',
          hclustfun=myclust,labRow=NA,
          main="Heatmap1",
          col = rev(brewer.pal(10,"RdYlBu"))
          )
```

```{r}
##testing
 clinic_data <- read.delim(file = "clinical_data.txt")
 state <- clinic_data[1,8]
 
##bind the results with original data
##here I use the codes of Rhodes Zhang
 gene_expression <- cbind(state,gene_expression)
 names(gene_expression)[which(names(gene_expression[,])=="PH-4")]=c("PH4")
 for(i in 1:3)
   {
       clinic_data$sampleID=sub("-",".",clinic_data$sampleID)
     }
 for( i in 1:length(gene_expression$state))
   {
       gene_expression[i,1]=clinic_data[which(clinic_data$sampleID==row.names(gene_expression[i,])),8]
 }
 
##count and print err rate 
 count <- 0
 result <- as.data.frame(result)
 for(i in 1:length(gene_expression[,1]))
 {
      if((as.character(gene_expression[i,1])=="Positive" && result[i,1]=="2") || (as.character(gene_expression[i,1])=="Negative" && result[i,1]=="1"))
          count=count+1
 }
```
##Error Rate

``` {r echo=FALSE}
 show("Error Rate:")
 show(count/length(gene_expression[,1]))
``` 
##Hierarchical Clustering Based on PCA data

```{r}
##load data
gene_expression <- read.table(file = "GeneMatrix.txt")
gene_expression <- t(gene_expression)
gene_expression <- as.data.frame(gene_expression)

##normalization
data <- scale(gene_expression)

##PCA
pca <- princomp(data,cor = T)

##get feature & choose principal component
feature <- pca$sdev^2
sum_f <- sum(feature)
k <- 389
while((sum(feature[1:k]) / sum_f) >= 0.75)
{
  k = k - 1
}

##get PCS & calculate cprs_data
pcs <- pca$loadings[,1:k]
sam <- length(data[,1])
cprs_data <- as.data.frame(0)
for(i in 1:k)
{
  for(j in 1:sam)
  {
    cprs_data[j,i] = matrix(data[j,],1,389) %*% matrix(pcs[,i],389,1)
  }
}

##Hierarchical Clustering & get the results of classification
rownames(cprs_data) <- rownames(gene_expression)
distance <- dist(cprs_data,method = "euclidean")
hc <- hclust(distance,method = "average")
d <- dist(gene_expression)
model <- hc
result=cutree(hc,k=2)


##the codes below is the same as former
myclust <- function(x){
  hclust(x , method = "average")
}
```
##Heatmap2
```{r echo=FALSE}
heatmap.2(as.matrix(d),
          trace='none',
          hclustfun=myclust,labRow=NA,
          main="Heatmap2",
          col = rev(brewer.pal(10,"RdYlBu"))
)
```

```{r}
clinic_data <- read.delim(file = "clinical_data.txt")
state <- clinic_data[1,8]



cprs_data <- cbind(state,cprs_data)
names(cprs_data)[which(names(cprs_data[,])=="PH-4")]=c("PH4")
for(i in 1:3)
{
  clinic_data$sampleID=sub("-",".",clinic_data$sampleID)
}
for( i in 1:length(cprs_data$state))
{
  cprs_data[i,1]=clinic_data[which(clinic_data$sampleID==row.names(cprs_data[i,])),8]
}



count <- 0
result <- as.data.frame(result)
for(i in 1:length(cprs_data[,1]))
{
  if((as.character(cprs_data[i,1])=="Positive" && result[i,1]=="2") || (as.character(cprs_data[i,1])=="Negative" && result[i,1]=="1"))
    count=count+1
}
```
##Error Rate

```{r echo=FALSE}
show("Error Rate:")
show(count/length(cprs_data[,1]))
```

##Discussion
In hc based on original data, the error rate is 6.32%. 
In hc based on PCA, because I set the threshold as 0.75 for I think the number of components should not be more than 50, which means components whose sum of scales is less than 25% are dropped, the error rate reduces to 9.58%. 
If I set the threshold of PCA as 0.8 or more, the result can be the same as hc based on original data, even better.
